Transparent Standby for Low-Power, Resource-Constrained Embedded Systems

### 2018-01-30 LCTES/body.tex
(>) LCTES

 Thank you, I will improve the page.

@@

In comparison to promises/futures, it provides lexical parallel constructs, allowing the branches to share local variables and, more importantly, supporting safe abortion of code (with the "par/or").

- application on IoT
- gains??

- structured programming
- lexical scope
- automatic power management: standby as well as enabled/disabled, powered, switched

needs a requires a energy-aware driver

process for ISR bindings and
A infra-structure programmer writes device drivers for
Each peripheral requires a driver
An architecture
Infra
Once a driver is written, all applications

- synchronous languages / \CEU

- except drivers, which are also written in Ceu

I have been working in the design of \CEU, a new reactive programming language
targeting resource-constrained embedded systems, for the past 8
years~\cite{ceu.sensys11,ceu.tr,ceu.sensys13,ceu.rem13,ceu.phd,ceu.rebls14,
ceu.mod15,ceu.rebls15,ceu.terra,ceu.media.webmedia16,ceu.tecs17}.
%
\CEU is grounded on the synchronous concurrency model, which has been
successfully adopted in the context of hard real-time systems such as avionics
and automobiles industry since the 80's~\cite{rp.twelve}.
%
The synchronous model trades power for reliability and has a simpler model
of time that suits most requirements of IoT applications.
%
On the one hand, this model cannot directly express time-consuming
computations, such as compression and cryptography algorithms, which are
typically either absent or delegated to auxiliary chips in the context of the
IoT.
%
On the other hand, all reactions to the external world are guaranteed to be
computed in bounded
time, ensuring that applications always reach an idle state amenable to standby
mode.

- results
    - atmega 328p/2560
    - arm cortex-m0
    - how much efficiency?
%
We expect that existing energy-unaware applications will benefit from savings
in the order of 50\% based on IEA's estimates considering current standby
technologies~\cite{iea.data}, and previous third-party work on transparent energy
awareness~\cite{wsn.tos.2}.

In Section~\ref{sec.ceu}, we introduce the synchronous model together with the
programming language \CEU.
In Section~\ref{sec.ext}, we present the language extensions that support
transparent standby.
In Section~\ref{sec.eval}, we evaluate .

\section{The Synchronous Language CEU}
\label{sec.ceu}

>>>>

- automatic power management: standby as well as enabled/disabled, powered, switched

- application on IoT

- results
    - atmega 328p/2560
    - arm cortex-m0
    - how much efficiency?

In Section~\ref{sec.ceu}, we introduce the structured synchronous model and the
programming language \CEU.
In Section~\ref{sec.ext}, we present the language extensions that support
transparent standby.
In Section~\ref{sec.eval}, we evaluate .

\section{The Structured Synchronous Programming Language \CEU}

\begin{comment}
- structured programming
- lexical scope

@@

In comparison to promises/futures, it provides lexical parallel constructs, allowing the branches to share local variables and, more importantly, supporting safe abortion of code (with the "par/or").
\end{comment}

\CEU is a Esterel-based\cite{ceu.tecs17} reactive programming language
targeting resource-constrained embedded systems~\cite{ceu.sensys13}.
%
It is grounded on the synchronous concurrency model, which has been
successfully adopted in the context of hard real-time systems such as avionics
and automobiles industry since the 80's~\cite{rp.twelve}.
%
The synchronous model trades power for reliability and has a simpler model
of time that suits most requirements of IoT applications.
%
On the one hand, this model cannot directly express time-consuming
computations, such as compression and cryptography algorithms, which are
typically either absent or delegated to auxiliary chips in the context of the
IoT.
%
On the other hand, all reactions to the external world are guaranteed to be
computed in bounded time, ensuring that applications always reach an idle state
amenable to standby mode.

\subsection{A Motivating Example}
\label{sec.ceu}

{\linespread{1}
\begin{figure}[t]
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[xrightmargin=1cm]
digitalWrite(13, 0);
int old = 0;
while (1) {
    int cur = digitalRead(2);
    if (cur != old) {
        digitalWrite(13, cur);
        old = cur;
    }
}

\end{lstlisting}
\centering\small{[a] Active polling in C}
\end{minipage}
%
\begin{minipage}[t]{0.45\linewidth}
\begin{lstlisting}[numbers=left]
input  int PIN_02;
output int PIN_13;

emit PIN_13(0);
loop do
    var int v = await PIN_02;
    emit PIN_13(v);
end
.
\end{lstlisting}
\centering\small{[b] Reactive execution in \CEU}
\end{minipage}
%\rule{8.4cm}{0.37pt}
\caption{ Controlling a LED from button state changes.
\label{lst.basic}
}
\end{figure}
}

Figure~\ref{lst.basic} shows a basic program that makes a button connected
to the microcontroller's pin 2 to control a LED connected to pin 13.
Whenever the state of the button changes (i.e., pressed or unpressed), the
state of the LED also changes (i.e., on or off).

The implementation in C in Figure~\ref{lst.basic}.a first turns off the LED
(ln. 1), then sets an initial state for the button (ln. 2), and then reads in a
loop (ln. 3-9) its current state (ln. 4) as fast as possible.
If the current state differs from the old state (ln. 5-8), the new state is
copied to the LED (ln. 6), and is also set as the old state (ln. 7).
%
This approach which actively samples the state of an external device (e.g., a
button) is known as \emph{polling} or \emph{software-driven I/O}.
%
On the one hand, this style does not involve dealing with the complexity of
callbacks (as discussed in Section~\ref{sec.related}) and is widely adopted in
Arduino, the most popular embedded platform (to be discussed in
Section~\ref{sec.method.software}).
On the other hand, polling wastes CPU cycles and prevents the device to enter in
standby.

The implementation in \CEU in Figure~\ref{lst.basic}.b
uses dedicated vocabulary to handle events from I/O devices without resorting
to callbacks.
The pins are first declared to determine their purpose in the program, i.e.,
\code{input} for the pin connected to the button and  \code{output} for the pin
connected to the LED (ln 1-2).
Then, the LED is turned off with an \code{emit} (ln. 4), which indicates an
output operation.
Then, the program enters in a loop to react to changes in the button (ln. 5-8).
The \code{await} primitive accounts for the reactive nature of \CEU.
The program literally waits for a change in the button (ln. 6) to be able to
proceed and set the new state of the LED (ln 7).
%
This way, when an application reaches an idle state, the language has precise
information about which events can awake that application.

By design, all lines of execution in \CEU always reach an \code{await} at the
end of a reaction to an event (otherwise, the application does not
compile~\cite{ceu.sensys13}).
%
This not only allows the application to enter standby mode, but effectively use
the deepest sleeping level considering all possible awaking events.
While in standby, only a hardware interrupt associated with the events can
awake the application, making it sleep for longest possible periods of time.
%
%This is still a theoretical assumption that needs to be proven in practice.

### 2018-01-30 LCTES/lctes.tex
(>) LCTES

\usepackage{booktabs} % For formal tables

@@

\begin{document}
\title{SIG Proceedings Paper in LaTeX Format}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}

@@

\begin{teaserfigure}
  %\includegraphics[width=\textwidth]{sampleteaser}
  \caption{This is a teaser}
  \label{fig:teaser}
\end{teaserfigure}

>>>>

\usepackage{booktabs} % For formal tables
\usepackage{xspace}

\usepackage{listings}
\usepackage{color}
    \definecolor{light}{gray}{0.97}
    \definecolor{dark}{gray}{0.30}
\lstset{
%columns=fullflexible,
%basicstyle=\ttfamily,
escapeinside={||},
    %mathescape=true,
    language=C, % choose the language of the code
    basicstyle=\fontfamily{pcr}\selectfont\scriptsize\color{black},
    keywordstyle=\color{black}\bfseries, % style for keywords
    numbers=none, % where to put the line-numbers
    numberstyle=\tiny, % the size of the fonts that are used for the line-numbers
    backgroundcolor=\color{light},
    showspaces=false, % show spaces adding particular underscores
    showstringspaces=false, % underline spaces within strings
    showtabs=false, % show tabs within strings adding particular underscores
    %frame=single, % adds a frame around the code
    tabsize=2, % sets default tabsize to 2 spaces
    %rulesepcolor=\color{gray}
    captionpos=b, % sets the caption-position to bottom
    breaklines=false, % sets automatic line breaking
    %breakatwhitespace=false,
    numbersep=2em,
    % C was used in the blocksworld example to refer to block C and nowhere else
    emph={par,or,hor,do,end,loop,code,await,pause,emit,input,event,call,with,%
          var,and,then,else,return,pure,deterministic,nohold,finalize,%
          class, every, FOREVER, this, spawn, in, pool, watching, until, 
          interface, each, abort, when, signal, PROC, CHAN, SIGNAL, PAR, not,
          bool, tag, escape, traverse,implementation,output,true,false,
          native,@const,@pure,@safe,define,public,private,none},
    emphstyle={\bfseries},
    commentstyle=\color{dark}\scriptsize,
    %xleftmargin=20pt,
    %xrightmargin=20pt,
    framesep=20pt,
    %upquote=true,
    %aboveskip={1.5\baselineskip},
}

\newcommand{\CEU}{\textsc{C\'{e}u}\xspace}
\newcommand{\code}[1] {{\small{\texttt{#1}}}}

@@

\begin{document}
%\title{A Language-Based Approach for Standby in Embedded Systems}
\title{Transparent Standby for Low-Power, Resource-Constrained Embedded Systems}
\subtitle{A Programming Language-Based Approach}

%\titlenote{Produces the permission block, and copyright information}
%\subtitlenote{The full version of the author's guide is available as \texttt{acmart.pdf} document}

@@

%\begin{teaserfigure}
  %\includegraphics[width=\textwidth]{sampleteaser}
  %\caption{This is a teaser}
  %\label{fig:teaser}
%\end{teaserfigure}
